---
title: <center>White-throated Sparrow Tarsus Length</center>
author: <center>Daniel Redwine</center>
date: <center>`r Sys.Date()`</center>
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 1  # up to one depth of headings, no subheader (specified by # in html)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme
---
```{r setup, include=FALSE}

# Load all libraries needed
library(ggplot2)
library(dplyr)
library(lme4) #GLM packagees 
library(lmerTest)
library(ggfortify)
library(MuMIn)
library(asbio)
library(GGally)
library(performance)
library(patchwork) # Easy way to arrange ggplot
library(olsrr) # Easy way to look at regression plots
library(RColorBrewer) # colors are fun 
library(emmeans)
library(tidyverse)

```

# White-throated Sparrow Tarsus

We are unsure whether tarsus length is a reliable measure of body size. If tarsus length is functionally equivalent to a measurement such as wing chord, then it does not make sense to continue measuring it in the field. Here we will investigate how measurements relate to tarsus length. <br>

```{r Load and clean dataset, echo=FALSE, message=FALSE, warning=FALSE}

wtsp_data <- read.csv("../data/wtsp_data_23_oct_23.csv") 
#load in dataset, ../ moves up one folder

# Select the variables I need and clean up dataset 
wtsp_data_clean <- wtsp_data %>% # name new object
  filter(New_Recap == "N") %>% # filter to only include new birds to avoid repeated measure
  # If some recaps didn't have tarsus initially may want to include
  select(SampleID, PCRsex, Wing, BirdMass, Tarsus) %>% # Select variables 
  filter(PCRsex == "M" | PCRsex == "F" ) %>% # make sure only M/F
  filter(Tarsus != "NA") # filter out any blanks
  
# Ensure that all variables are factors/numeric as appropriate
  wtsp_data_clean$SampleID <- as.factor(wtsp_data_clean$SampleID)
  wtsp_data_clean$PCRsex <- as.factor(wtsp_data_clean$PCRsex)
  wtsp_data_clean$Wing <- as.numeric(wtsp_data_clean$Wing)
  wtsp_data_clean$BirdMass <- as.numeric(wtsp_data_clean$BirdMass)
  wtsp_data_clean$Tarsus <- as.numeric(wtsp_data_clean$Tarsus)
  
```

# Visualizing data

We'll look at some plots of wing chord, sex, and mass vs tarsus as well as sex vs wing chord length. Missing values for morph pcr likely precludes looking at how morph relates to tarsus length for now. <br>

```{r tarsus mass and wing plot, echo=FALSE, message=FALSE, warning=FALSE, fig.show = "hold", fig.height = 7, out.width = "90%", fig.align = "center", fig.cap="Figure 1. Graphs of wing chord length (mm), mass (g), and sex vs tarsus length (mm) as well as sex vs wing chord length (mm)."}


wing_tarsus_line <- ggplot(wtsp_data_clean, aes(Wing, Tarsus)) + 
  geom_point() +
  geom_smooth(method="lm") +
  theme_bw() +
  ylab("Tarsus Length (mm)") +
  xlab("Wing Chord Length (mm)")

ggsave("wing_vs_tarsus_plot.png", path ="../script/output") #../ moves up a level

mass_tarsus_line <- ggplot(wtsp_data_clean, aes(BirdMass, Tarsus)) + 
  geom_point() +
  geom_smooth(method="lm") +
  theme_bw() +
  ylab("Tarsus Length (mm)") +
  xlab("Mass (g)")

ggsave("mass_vs_tarsus_plot.png", path ="../script/output") #../ moves up a level

sex_tarsus_box <- ggplot(wtsp_data_clean, aes(PCRsex, Tarsus, fill = PCRsex)) + 
  geom_boxplot() +
  theme_bw() +
  xlab("Sex") +
  ylab("Tarsus Length (mm)") +
  theme(legend.position  = "none") +
  scale_fill_brewer(palette = "Blues")

ggsave("sex_vs_tarsus_boxplot.png", path ="../script/output") #../ moves up a level

sex_wing_box <- ggplot(wtsp_data_clean, aes(PCRsex, Wing, fill = PCRsex)) + 
  geom_boxplot() +
  theme_bw() +
  xlab("Sex") +
  ylab("Wing Chord Length (mm)") +
  theme(legend.position  = "none") +
  scale_fill_brewer(palette = "Purples")

ggsave("sex_vs_wing_boxplot.png", path ="../script/output") # up one level works for Rmd and script


wing_tarsus_line / sex_tarsus_box | mass_tarsus_line / sex_wing_box
# patchwork notation for arranging plots
# | puts plots next to eachother while / puts plots over another
# wing tarsus line over sex tarsus box next to mass tarsus line over sex wing box
```
<br>
There may be a correlation of both mass and wing chord length to tarsus length. However, this may not be statistically significant. There appears to be a greater difference between wing chord lengths than tarsus length between sexes. Looks like there are a couple of outlier points for tarsus length. <br>

# Data Analysis

**Tarsus (mm) as a function of wing chord length (mm)**

```{r wing tarsus model, message=FALSE, echo=FALSE, warning=FALSE}
w_t_model <- lm(Tarsus ~ Wing, data=wtsp_data_clean) # make the model 
summary(w_t_model) # model summary 
coef(w_t_model) # coefficients for slope and r squared
```
<br>
**Tarsus = 0.11 x Wing  + 16**
<br><br>
**R^2 = 0.09376**
<br><br>
**p < 0.01**
<br>
```{r check wing tarsus model, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.show = "hold", out.width = "90%", fig.align = "center", fig.cap="Figure 2. Assumptions for the model predicting tarsus by wing chord lenth. Looks good."}
performance::check_model(w_t_model) # checking model assumptions
```
